PTFO 8-8695: Two Stars, Two Signals, No Planet

### 2020-04-27 paper/ms.tex
JNW edits
Co-authored-by: joshwinn <jwinn@astro.princeton.edu>

This also merges https://github.com/lgbouma/billy/pull/2

%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: Still Against the Planetary Interpretation of
% PTFO$\,$8-8695b}
% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
% Planet}
%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
%PTFO$\,$8-8695b is Not a Planet}
%\title{Double Trouble: Transient Dips and Photometric Binarity in
%PTFO$\,$8-8695}
\title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
PTFO 8-8695}
\title{PTFO\,8-8695: A binary star, not a hot Jupiter host}

\correspondingauthor{L. G. Bouma}
\email{luke@astro.princeton.edu}

%
% key authors:
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J.\,N.\,Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\begin{abstract}
  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year
  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
  While strong evidence against the planet hypothesis has already been provided,
  the possibility remains debated in the literature.
  To obtain further clues, we inspected data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and independently suggest that it is an
  astrometric binary.  The TESS lightcurve shows two different photometric
  periods. The variability is dominated by a sinusoidal signal with a period
  of 11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
  by 1.2\% hour-long dips, the type of signal previously
  intepreted as planetary transits.
  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
  which, as noted previously, makes them difficult to explain as planetary transits.
  Instead, we believe that PTFO$\,$8-8695 is a pair of young
  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
  behavior that been seen in at least 5 other cases.
  The origin of these transient dips is still unknown but likely involves circumstellar material.
\end{abstract}

@@

\epigraph{
	Nubble [n]: a small lump or protuberance.
}

@@

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two different periodic signals, which we analyze in
Section~\ref{sec:tess}.
The Gaia data, analyzed in Section~\ref{sec:gaia}, show that
\ptfo\ is a photometric binary, and suggest it also an astrometric binary.
We discuss the pieces of the puzzle
in Section~\ref{sec:discussion}, and summarize the situation in
Section~\ref{sec:conclusions}.
% JNW: this kind of stuff is for the abstract/conclusion, not intro.
%The phase of the dip within the seems
%to have instantaneously changed over the past decade.  In addition,
%other young stars show lightcurve morphologies similar to the
%short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
%binary M dwarf in which one star shows a rotation signal, and the
%other shows transient dipping potentially caused by eclipses of
%lasma at the Keplerian co-rotation radius.

@@

% JNW: How about more descriptive labels? f --> Total flux
% fl = Longer-period signal
% fs = Shorter-period signal
% last panel, Residuals

@@

\ptfo\ [also known as CVSO\,30; \cite{briceno_cida_2005}}]
was observed by TESS with Camera 1, CCD 1, from December 15,
2018 until January 6, 2019, during the sixth sector of science operations
\citep{ricker_transiting_2015}.  The star was designated TIC\,264461976
in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
The pixel data for an $11\times11$ array surrounding \ptfo\ were
averaged into 2-minute stacks by the onboard computer.  Each
2048$\times$2048 image from the CCD was also averaged into 30-minute
stacks, and saved as a ``full frame image'' (FFI).

@@

As an independent check on the 2-minute SPOC lightcurve, we
examined the lightcurve based upon 30-minute image stacks which was
produced as part of the Cluster
Difference Imaging Photometric Survey (CDIPS;
\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve of choice used a
circular aperture with radius 1 pixel.
% JNW: What does that mean? You used fractional pixels in calculating the pixel sums?

@@

We normalized the lightcurve by dividing out the median flux, and then opted
to subtract 1.0 to set the median value to zero, which simplified
subsequent interpretation.  Many
of these and subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}.

@@

Between July 25, 2014 and May 23, 2016, Gaia measured about 300
billion centroid positions of 1{.}6 billion stars
\citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
In the Gaia second data release (DR2), these CCD observations were
used to determine positions, proper motions, and parallaxes of the
brighest 1{.}3 billion stars, including \ptfo\
\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good'' observations, i.e., observations that were not strongly down-weighted in
the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
3222255959210123904.  It photometric brightness was measured using
selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
the pipeline parameters for \ptfo\ using the Gaia
archive\footnote{\url{gea.esac.esa.int/archive/}}.

@@

With the Gaia astrometry, it has become clear that 25$\,$Ori itself
has distinct subgroups
\citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
cluster membership of \ptfo, we follow the notation and results of
\citet{kounkel_apogee2_2018}.  These authors combined astrometric data
from Gaia DR2 with near-infrared spectra from APOGEE-2
\citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
They performed a hierarchical clustering on the six-dimensional
position and velocity information to identify subgroups within the
Orion complex.  From smallest to largest, \ptfo\ was identified
as a member of the following hierarchical subgroups:
\begin{equation}
  %{\rm PTFO\,8\text{-}8695}
  % \in
  {\rm 25\,Ori\text{-}1}
  \subset {\rm 25\,Ori}
  \subset {\rm Orion\ OB1a}
  \subset {\rm Orion\ D},
\end{equation}
where `$\subset$' means ``is a proper subset
of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
identified members.  Based on the CMD, the mean age was found to be
6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
\citep{kounkel_apogee2_2018}. 
% JNW: What's the difference between a CMD and an HR diagram?
\citet{kounkel_apogee2_2018} identified
seven other smaller groups in the Orion complex near the Be star
25$\,$Ori. These groups received higher numbers, {\it e.g.},
25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).

@@

% JNW This doesn't seem necessary to me.
%Initial signal-processing experiments fitting out splines or sinusoids
%showed that after subtracting out the long-period signal, the
%short-period signal dominated the periodogram, and vice-versa.
%However it quickly became clear that it would be beneficial to
%simultaneously model each signal, in order to preserve the
%power at each frequency.

@@

We fitted a model to the lightcurve consisting of a linear combination of Fourier modes
with periods $P_{\rm s}$ and $P_{\rm ell}$, as well as a number of harmonics chosen as described below. To try accounting for the dips, we also added an
analytic transit model with period $P_{\rm s}$.
Symbolically, the total flux $f$ is given as
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell}
  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the flux at the shorter period, and
$f_{\rm \ell}$ is the flux at the longer period.  Writing out the
Fourier terms explicitly,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{N} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{N} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{M} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{M} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
where $N$ and $M$ are the total number of modes at the shorter and longer
periods, respectively,
% JNW: maybe N_\ell and N_s?
$A_i$ and $B_i$ the amplitudes of each
mode (which can be positive or negative), and $\omega_\ell$
and $\omega_{\rm s}$ are the angular frequencies of the longer-period
and shorter-period signals. By not including a phase parameter
in the shorter-period model, we have implicitly defined the zero point
of the phase scale. The relative phase of the longer-period model
is specified by the phase parameter $\phi_\ell$.
Since we did not know in advance how many harmonics
would be appropriate to include in the model, we considered a number of different choices for
$N$ and $M$, and used the Bayesian information criterion to select
the final model (Table~1).

As an example, one possible model consists of a transit, $N=2$
sines and cosines at the shorter period, plus $M=1$
sine and cosine at the longer period.  In this case, the free parameters would
be as follows.  The transit model parameters are the impact
parameter, the planet-to-star radius ratio, two quadratic limb
darkening parameters, the planet's orbital period (set equal to $P_{\rm s}$)
the time of a particular transit, and the mean flux.  There
would be $2N=4$ additional Fourier amplitudes at the shorter period,
plus $2M=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
itself and the relative phase $\phi_\ell$.  The total number
of parameters is 14.

We implemented and fitted the models using \texttt{PyMC3}, which is
built on \texttt{theano}
\citep{salvatier_2016_PyMC3,exoplanet:theano}.
For the Fourier terms,
we used the default math operators.
% JNW what does that mean? If you just mean 'numpy', is it necessary to state?
For the exoplanet transit, we
used the model and derivatives implemented in the \texttt{exoplanet} code
\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
speed up the fitting process, we averaged the 2-minute lightcurve
to 10-minute samples.  We correspondingly scaled the uncertainties in
the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
initialized each model with the parameters of the maximum a posteriori (MAP) model.
We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
diagnostic \citep{gelman_inference_1992}.  We tested our ability to
successfully recover injected parameters using synthetic data before
fitting the \ptfo\ lightcurves.

@@

We considered nine models, with the number of modes per frequency
$N$ and $M$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion (Table~1).  The model with
the lowest BIC had two modes at the shorter 10.74$\,$hr period, and
two modes at the longer 11.96$\,$hr period.  The next-best model had
an additional harmonic at the longer period ($M=3$), but was otherwise
identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
and 1.51, which suggests a plausible though imperfect agreement
between the data and the model to within the formal uncertainties.
Table~2 gives the best-fitting parameters for the preferred model,
which has the lowest BIC value.

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene used for blend analysis.}
		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
		logarithnic grayscale. The yellow star is the position of \ptfo\.
		Orange crosses are neighboring stars with $T<17$. The
		\texttt{X} and \texttt{/} hatches show the apertures
		used to measure the background and target star flux, respectively.
		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
		field, with a linear grayscale. The circles show the apertures of radii
		1, 1.5, and 2.25 pixels used in our blend analysis. To the northwest
		of \ptfo\ and between the
		blue and orange circles is ``Star A'',
		the only star bright and close enough to be contributing
		to the signal attributed to \ptfo\. However, the pixel-level TESS data
		showed that any such contamination is negligible (see
		Section~\ref{subsec:blend}).
		% JNW Label Star A?
		\label{fig:scene}
	}
\end{figure}

@@

%FIXME: would be better to actually be able to get them...
While long-baseline radial velocity (RV) measurements could provide
complementary lower limits on companion masses and separations, the available RV
data for \ptfo\ have large uncertainties due to the faintness of the star.  The
RV dataset with the longest timespan in the literature appears to be the 5
Keck/HIRES measurements acquired over 10 days in April 2011 by
\citet{van_eyken_ptf_2012}.  The root-mean-squared RV over that 10 day span was
$1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with the measurement
precision.  Although \citet{yu_tests_2015} acquired 22 further
Keck/HIRES spectra on one night in December 2013, those points
were not reduced to velocities. Further Keck/HIRES
measurements of this target could confirm or refute the existence of
possible binary companions.

@@

%Given our detection of two separate signals, whether \ptfo\ could be a
%photometric binary was of great interest.  
Figure~\ref{fig:gaia} shows
the HR diagram we constructed from the data.  The diagram shows
that \ptfo\ is $\approx$0.75 magnitudes brighter than the average
25$\,$Ori-1 star of the same color.  In other words, it is about twice
as bright as expected for a single star in the cluster.
It also seems to be part of a ``photometric binary''
track that runs parallel to the main track.

@@

In their report on the discovery of the unusual
photometric variability, \citet{van_eyken_ptf_2012} saw an alias of the
same signal ({\it e.g.}, their Figure~7),
% JNW: Don't understand this sentence. What is the 'same signal'?
and identified it as a
periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
their observing cadence, because of its close correspondence to the
sidereal day.
% JNW What is this next phrase doing here?
%While the TESS data can show significant reflected
%light from the Earth \citep[{\it e.g.},][]{luger_tess_2019}, 
Our
pixel-level analysis showed that the signal is specific to only pixels
near \ptfo, and no other pixels.  We therefore conclude that the
signal is not an artifact of systematic errors.

@@

Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
s}$.  The observation of multiple dips per cycle in 2015 was seen
independently by both \citet{yu_tests_2015} and
\citet{tanimoto_evidence_2020}.  It therefore seems credible.
Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
seems less plausible---the phase $-0{.}47$ dips are strongly detected,
while the suggested phase 0 dip is not clearly present in the data.

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
In the terminology of \citet{stauffer_orbiting_2017},
these objects are morphologically distinct from ``scallop shell'' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars all show angular dips
that are cannot be explained as the effects of starspots.
These stars typically have the following things in common:
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item They are weak-lined T Tauri stars.
  \item The spectral type is M2 to M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item The age is $\lesssim$ 100$\,$Myr.
%  \item They lack spectroscopic indicators of strong accretion.
% JNW: Same as WTTS?
  \item The lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item The rotation is rapid---usually between 0.5 and 2.0 days.
  \item There is rarely any detectable infrared excess in WISE data
  (never any W4 detection; only a few W3 detections).
  \item They sometimes show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item The dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item The dip depths can change after flares.
  \item They are rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

% \clearpage
\input{model_comparison_table.tex}
\input{model_posterior_table.tex}
% \clearpage

>>>>

%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: Still Against the Planetary Interpretation of
% PTFO$\,$8-8695b}
% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
% Planet}
%\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
%PTFO$\,$8-8695b is Not a Planet}
%\title{Double Trouble: Transient Dips and Photometric Binarity in
%PTFO$\,$8-8695}
% \title{Double, Double, Spin plus Nubble: Two Stars and Two Signals in
% PTFO 8-8695}
% \title{PTFO\,8-8695: A binary star, not a hot Jupiter host}   
%\title{Two Stars and Two Signals in PTFO 8-8695}
\title{PTFO 8-8695: Two Stars, Two Signals, No Planet}

\correspondingauthor{L.\,G.\,Bouma}
\email{luke@astro.princeton.edu}

%
% key authors:
%
\author[0000-0002-0514-5538]{L.\,G.\,Bouma}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{J.\,N.\,Winn}
\affiliation{ Department of Astrophysical Sciences, Princeton
University, 4 Ivy Lane, Princeton, NJ 08540, USA}
%
\author[0000-0002-4265-047X]{TESS collaborators}
\affiliation{ Department of Astrophysical Sciences, Princeton
	University, 4 Ivy Lane, Princeton, NJ 08540, USA}

\begin{abstract}
  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year
  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
  While strong evidence against the planet hypothesis has already been provided,
  the possibility remains debated in the literature.
  To obtain further clues, we inspected data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and independently suggest that it is an
  astrometric binary.  The TESS lightcurve shows two different photometric
  periods. The variability is dominated by a sinusoidal signal with a period
  of 11.96$\,$hr, presumably caused by stellar rotation.  Also
  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
  by hour-long dips, the type of signal previously
  intepreted as planetary transits.
  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
  which, as noted previously, makes them difficult to explain as planetary transits.
  Instead, we believe that PTFO$\,$8-8695 is a pair of young
  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
  behavior that been seen in at least 5 other cases.
  The origin of these transient dips is still unknown but likely involves circumstellar material.
\end{abstract}

@@

@@

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two different periodic signals, which we analyze in
Section~\ref{sec:tess}.
The Gaia data, analyzed in Section~\ref{sec:gaia}, show that
\ptfo\ is a photometric binary, and suggest it also an astrometric binary.
We discuss the pieces of the puzzle
in Section~\ref{sec:discussion}, and summarize the situation in
Section~\ref{sec:conclusions}.

@@

@@

\ptfo\ [also known as CVSO\,30; \citealt{briceno_cida_2005}]
was observed by TESS with Camera 1, CCD 1, from December 15,
2018 until January 6, 2019, during the sixth sector of science operations
\citep{ricker_transiting_2015}.  The star was designated TIC\,264461976
in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
The pixel data for an $11\times11$ array surrounding \ptfo\ were
averaged into 2-minute stacks by the onboard computer.  Each
2048$\times$2048 image from the CCD was also averaged into 30-minute
stacks, and saved as a ``full frame image'' (FFI).

@@

As an independent check on the 2-minute SPOC lightcurve, we
examined the lightcurve based upon 30-minute image stacks which was
produced as part of the Cluster
Difference Imaging Photometric Survey (CDIPS;
\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve of choice used a
circular aperture with radius 1 pixel.

@@

We normalized the lightcurve by dividing out the median flux, and then opted
to subtract 1{.}0 to set the median value to zero, which simplified
subsequent interpretation.  Many
of these and subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}.

@@

Between July 25, 2014 and May 23, 2016, Gaia measured about 300
billion centroid positions of 1{.}6 billion stars
\citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
In the Gaia second data release (DR2), these CCD observations were
used to determine positions, proper motions, and parallaxes of the
brighest 1{.}3 billion stars, including \ptfo\
\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good''
observations, {\it i.e.}, observations that were not strongly down-weighted in
the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
3222255959210123904.  It photometric brightness was measured using
selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
the pipeline parameters for \ptfo\ using the Gaia
archive\footnote{\url{gea.esac.esa.int/archive/}}.

@@

With the Gaia astrometry, it has become clear that 25$\,$Ori itself
has distinct subgroups
\citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
cluster membership of \ptfo, we follow the notation and results of
\citet{kounkel_apogee2_2018}.  These authors combined astrometric data
from Gaia DR2 with near-infrared spectra from APOGEE-2
\citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
They performed a hierarchical clustering on the six-dimensional
position and velocity information to identify subgroups within the
Orion complex.  From smallest to largest, \ptfo\ was identified
as a member of the following hierarchical subgroups:
\begin{equation}
  %{\rm PTFO\,8\text{-}8695}
  % \in
  {\rm 25\,Ori\text{-}1}
  \subset {\rm 25\,Ori}
  \subset {\rm Orion\ OB1a}
  \subset {\rm Orion\ D},
\end{equation}
where `$\subset$' means ``is a proper subset
of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
identified members.  Based on the CMD, the mean age was found to be
6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
\citep[see][Section 2.3]{kounkel_apogee2_2018}. 
\citet{kounkel_apogee2_2018} identified
seven other smaller groups in the Orion complex near the Be star
25$\,$Ori. These groups received higher numbers, {\it e.g.},
25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).

@@

@@

We fitted a model to the lightcurve consisting of a linear combination of Fourier modes
with periods $P_{\rm s}$ and $P_{\rm ell}$, as well as a number of harmonics chosen as described below. To try accounting for the dips, we also added an
analytic transit model with period $P_{\rm s}$.
Symbolically, the total flux $f$ is given as
\begin{equation}
  f = f_{\rm s} + f_{\rm \ell}
  = f_{\rm transit,s} + f_{\rm Fourier,s} + f_{\rm Fourier,\ell},
\end{equation}
where $f_{\rm s}$ is the flux at the shorter period, and
$f_{\rm \ell}$ is the flux at the longer period.  Writing out the
Fourier terms explicitly,
\begin{align}
  f = &f_{\rm transit,s} + \sum_{n=1}^{N_{\rm s}} A_n \sin(n\omega_{\rm s}t)
  + \sum_{n=1}^{N_{\rm s}} B_n \cos(n\omega_{\rm s}t)\\
  &+ \sum_{m=1}^{N_{\rm \ell}} A_m \sin(m[\omega_{\rm \ell}t+\phi_{\rm \ell}])
  + \sum_{m=1}^{N_{\rm \ell}} B_m \cos(m[\omega_{\rm \ell}t+\phi_{\rm \ell}]), \nonumber
\end{align}
where $N_{\rm s}$ and $N_{\rm \ell}$ are the total number of modes at the shorter and longer
periods, respectively,
$A_i$ and $B_i$ are the amplitudes of each
mode (which can be positive or negative), and $\omega_\ell$
and $\omega_{\rm s}$ are the angular frequencies of the longer-period
and shorter-period signals. By not including a phase parameter
in the shorter-period model, we have implicitly defined the zero point
of the phase scale. The relative phase of the longer-period model
is specified by the phase parameter $\phi_\ell$.
Since we did not know in advance how many harmonics
would be appropriate to include in the model, we considered a number of different choices for
$N_{\rm s}$ and $N_{\rm \ell}$, and used the Bayesian information criterion to select
the final model (Table~1).

As an example, one possible model consists of a transit, $N_{\rm s}=2$
sines and cosines at the shorter period, plus $N_{\rm \ell}=1$
sine and cosine at the longer period.  In this case, the free parameters would
be as follows.  The transit model parameters are the impact
parameter, the planet-to-star radius ratio, two quadratic limb
darkening parameters, the planet's orbital period (set equal to $P_{\rm s}$)
the time of a particular transit, and the mean flux.  There
would be $2N_{\rm s}=4$ additional Fourier amplitudes at the shorter period,
plus $2N_{\rm \ell}=2$ Fourier amplitudes at the longer period, and well as $P_\ell$
itself and the relative phase $\phi_\ell$.  The total number
of parameters is 14.

We implemented and fitted the models using \texttt{PyMC3}, which is
built on \texttt{theano}
\citep{salvatier_2016_PyMC3,exoplanet:theano}.
For the Fourier terms,
we used the default math operators.
For the exoplanet transit, we
used the model and derivatives implemented in the \texttt{exoplanet} code
\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
speed up the fitting process, we averaged the 2-minute lightcurve
to 10-minute samples.  We correspondingly scaled the uncertainties in
the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
initialized each model with the parameters of the maximum a posteriori (MAP) model.
We then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
diagnostic \citep{gelman_inference_1992}.  We tested our ability to
successfully recover injected parameters using synthetic data before
fitting the \ptfo\ lightcurves.

@@

We considered nine models, with the number of modes per frequency
$N_{\rm s}$ and $N_{\rm \ell}$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion (Table~1).  The model with
the lowest BIC had two modes at the shorter 10.74$\,$hr period, and
two modes at the longer 11.96$\,$hr period.  The next-best model had
an additional harmonic at the longer period ($N_{\rm \ell}=3$), but was otherwise
identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
and 1.51, which suggests a plausible though imperfect agreement
between the data and the model to within the formal uncertainties.
Table~2 gives the best-fitting parameters for the preferred model,
which has the lowest BIC value.

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Scene used for blend analysis.}
		{\it Top:} Mean TESS image of \ptfo\ over Sector~6, with a
		logarithmic grayscale. The yellow star is the position of \ptfo.
		Orange crosses are neighboring stars with $T<17$. The
		\texttt{X} and \texttt{/} hatches show the apertures
		used to measure the background and target star flux, respectively.
		{\it Bottom:} Digitized Sky Survey $R$-band image of the same
		field, with a linear grayscale. The circles show the apertures of radii
		1, 1.5, and 2.25 pixels used in our blend analysis. To the northwest
		of \ptfo\ and between the
		blue and orange circles is ``Star A'',
		the only star bright and close enough to be contributing
		to the signal attributed to \ptfo\. However, the pixel-level TESS data
		showed that any such contamination is negligible (see
		Section~\ref{subsec:blend}).
		\label{fig:scene}
	}
\end{figure}

@@

While long-baseline radial velocity (RV) measurements could provide
complementary lower limits on companion masses and separations, the available RV
data for \ptfo\ have large uncertainties due to the faintness of the star.  The
RV dataset with the longest timespan in the literature appears to be the 5
Keck/HIRES measurements acquired over 10 days in April 2011 by
\citet{van_eyken_ptf_2012}.  The root-mean-squared RV over that 10 day span was
$1.6\,{\rm km}\,{\rm s}^{-1}$, consistent with the measurement
precision.  Although \citet{yu_tests_2015} acquired 22 further
Keck/HIRES spectra on one night in December 2013, those points
were not reduced to velocities. Further Keck/HIRES
measurements of this target could confirm or refute the existence of
possible binary companions.

@@

Figure~\ref{fig:gaia} shows
the HR diagram we constructed from the data.  The diagram shows
that \ptfo\ is $\approx$0.75 magnitudes brighter than the average
25$\,$Ori-1 star of the same color.  In other words, it is about twice
as bright as expected for a single star in the cluster.
It also seems to be part of a ``photometric binary''
track that runs parallel to the main track.

@@

In their report on the discovery of the unusual
photometric variability, \citet{van_eyken_ptf_2012} saw an alias of the
longer-period signal ({\it e.g.}, their Figure~7),
and identified it as a
periodogram peak at $0.9985 \pm 0.0061\,$days. They ascribed it to
their observing cadence, because of its close correspondence to the
sidereal day.
Our
pixel-level analysis showed that the signal is specific to only pixels
near \ptfo, and no other pixels.  We therefore conclude that the
signal is not an artifact of systematic errors.

@@

Figure~\ref{fig:o_minus_c} shows two additional strange features: {\it
(i)}  multiple dips per cycle, and {\it (ii)} a set of dips
numerically coincident with phase $(P_{\rm \ell} - P_{\rm s}) / P_{\rm
s}$.  The observation of multiple dips per cycle in 2015 was seen
independently by both \citet{yu_tests_2015} and
\citet{tanimoto_evidence_2020}.  It therefore seems credible.
Inspecting the \citet{tanimoto_evidence_2020} lightcurves, the claim
of multiple dips per cycle in December 2018 at phase 0 and $-0{.}47$
seems less plausible---the phase $\text{-}0{.}47$ dips are strongly detected,
while the suggested phase 0 dip is not clearly present in the data.

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
In the terminology of \citet{stauffer_orbiting_2017},
these objects are morphologically distinct from ``scallop shell'' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars all show angular dips
that are cannot be explained as the effects of starspots.
These stars typically have the following things in common:
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item They are weak-lined T Tauri stars.
  \item The spectral type is M2 to M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item The age is $\lesssim$ 100$\,$Myr.
  \item The lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item The rotation is rapid---usually between 0.5 and 2.0 days.
  \item There is rarely any detectable infrared excess in WISE data
  (never any W4 detection; only a few W3 detections).
  \item They sometimes show multiple dips per cycle.
  \item The dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item The dip depths can change after flares.
  \item They are rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

\clearpage
\input{model_comparison_table.tex}
\input{model_posterior_table.tex}
\clearpage
